---
title: "Doublet identification report"
author: "SS"
date: "2023-12-05"
output: html_document
params: 
  data_object: "data_object"
  doublet_tbl: "doublet_tbl"
  annotation_tbl: "annotation_tbl"
  sample_names: "sample_names"
---
## Introduction 
This report contains UMAP representation of cell clusters and visualization of the distribution of doublets across processed samples.

```{r, include=FALSE}
library(purrr)
library(dplyr)
library(tidyr)
library(ggrepel)
library(Seurat)
library(glue)
library(scDblFinder)
library(tidyseurat)
library(tidySingleCellExperiment)
library(patchwork)
library(tibble)
library(scran)
library(magrittr)
sample_column <- "orig.ident" 
cell_ann_col <- "seurat_annotations" 
```

```{r, include=FALSE}
library(purrr)
calc_UMAP <- function(input_seurat) {
  assay_name <- input_seurat@assays |> names() |> extract2(1)
  
  # Check if variable features are already present, if not calculate them
  if (length(VariableFeatures(input_seurat)) == 0) {
    input_seurat <- FindVariableFeatures(input_seurat)
  }
  
  # Extract variable features using VariableFeatures() for Seurat v5
  var_genes <- VariableFeatures(input_seurat)
  
  # Ensure that there are variable features before proceeding
  if (length(var_genes) > 0) {
    # Scale data and run PCA on variable genes
    x <- ScaleData(input_seurat) |>
      RunPCA(features = var_genes) |>
      FindNeighbors(dims = 1:30) |>
      FindClusters(resolution = 0.5) |>
      RunUMAP(dims = 1:30, spread = 0.5, min.dist = 0.01, n.neighbors = 10L) |>
      as_tibble()
  } else {
    stop("No variable features available for UMAP calculation.")
  }
  
  return(x)
}

calc_UMAP_dbl_report <- map(data_object, calc_UMAP)

```

```{r, include=FALSE}
library(dplyr)
library(tidyr)
library(purrr)
# library(sccomp)
library(ggrepel)
library(Seurat)
library(glue)
library(scDblFinder)
library(Seurat)
library(tidyseurat)
library(tidySingleCellExperiment)
library(patchwork)
library(tibble)
library(scran)

get_labels_clusters = function(.data, label_column, dim1, dim2){

    tidy_dist = function(x1, x2, y1, y2){

        tibble(x1, x2, y1, y2) %>%
            rowwise() %>%
            mutate(dist = matrix(c(x1, x2, y1, y2), nrow = 2, byrow = T) %>% dist()) %>%
            pull(dist)

    }

    label_column = enquo(label_column)
    dim1 = enquo(dim1)
    dim2 = enquo(dim2)

    .data %>%
        nest(data = -!!label_column) %>%
        mutate(
            !!dim1 := map_dbl(data, ~ .x %>% pull(!!dim1) %>% median()),
            !!dim2 := map_dbl(data, ~ .x %>% pull(!!dim2) %>% median())
        ) %>%
        dplyr::select(-data)
}
```

## Comprehensive UMAP Visualization of Cell Typing and Doublet Detection Across Tissue Samples
- This visualization highlights the clustering of cell types and identifies singlets and doublets in the population. 
- This allows for an exploration of similarities and differences in gene expression profiles between cells from different tissues. 

<!-- ```{r, out.width='100%', fig.width=15, fig.height=10, warning=FALSE, message=FALSE, echo=FALSE} -->
<!-- ## Adjusting plot size according to the number of samples  -->
<!-- # num_samples <- length(params$x1$Tissue) -->
<!-- #  -->
<!-- # # Calculate the grid layout based on the number of samples -->
<!-- # num_columns <- ceiling(sqrt(num_samples)) -->
<!-- # num_rows <- ceiling(num_samples / num_columns) -->


<!-- # Joining info and returning a list opf tibbles -->
<!-- merged_combined_annotation_doublets <- list( -->
<!--   #params$x1, -->
<!--   calc_UMAP_dbl_report, -->
<!--   doublet_tbl, -->
<!--   annotation_tbl -->
<!-- ) |> -->
<!--     pmap( -->
<!--         ~ ..1 |> -->
<!--             left_join(..2, by = ".cell") |> -->
<!--             left_join(..3, by = ".cell")  -->
<!--           #left_join(..4, by = ".cell") -->
<!--     ) |>  -->
<!--   enframe(name = "sample_id", value = "annotated_metadata")|>  -->
<!--   mutate( -->
<!--     sample_name = sample_names) |>  -->
<!--   mutate(plot_by_doublet = map2( -->
<!--     annotated_metadata, -->
<!--     sample_name, -->
<!--           ~ { -->
<!--             #browser() -->
<!--             merged_combined_annotation_doublets = .x |> -->

<!--             #Sample to non overwhelmm the plotting -->
<!--             nest(doublet_class = -scDblFinder.class) |> -->
<!--             mutate(number_to_sample = if_else(scDblFinder.class=="singlet", 10000, Inf)) |> -->
<!--             replace_na(list(number_to_sample = Inf)) |> -->
<!--             # mutate( -->
<!--             #   doublet_class = map2(doublet_class, number_to_sample, ~ .x |> sample_n(min(n(), .y))) -->
<!--             #   ) -->
<!--             unnest(doublet_class)  -->
<!--             # doublet_plots <- plot_by_doublet$doublet_class[[1]]|> -->

<!--             merged_combined_annotation_doublets |>  -->
<!--             ggplot(aes(umap_1, umap_2, color = scDblFinder.class)) + -->
<!--               geom_point(shape=".", size = 10) + -->
<!--               theme_bw() + -->
<!--               labs(title = .y, color = "Cell Type") +  -->
<!--               ggrepel::geom_text_repel( -->
<!--                 data= get_labels_clusters( -->
<!--           			.x,  -->
<!--           			scDblFinder.class, -->
<!--           			umap_1,  -->
<!--           			umap_2 -->
<!--           			) , -->
<!--       		aes(umap_1, umap_2, label = scDblFinder.class), size = 4) + -->
<!--               guides(color = "none")+  -->
<!--         labs(title = .y) -->
<!--           #print(plot_by_doublet) -->
<!--            #return(plot_by_doublet) -->
<!--           })) |>  -->

<!--   mutate(plot_by_cell_type = map2( -->
<!--     annotated_metadata, -->
<!--     sample_name, -->
<!--           ~ { -->
<!--       #browser() -->
<!--       merged_combined_annotation_doublets = .x |> -->
<!--       # Sample to non overwhelmm the plotting -->
<!--       nest(doublet_class = -scDblFinder.class) |> -->
<!--       mutate(number_to_sample = if_else(scDblFinder.class=="singlet", 10000, Inf)) |> -->
<!--         replace_na(list(number_to_sample = Inf)) |> -->
<!--         #mutate(doublet_class = map2(doublet_class, number_to_sample, ~ .x |> sample_n(min(n(), .y))))|> -->
<!--       unnest(doublet_class) -->

<!--       merged_combined_annotation_doublets |>  -->
<!--       ggplot(aes(umap_1, umap_2, color = all_of(cell_ann_col))) + -->
<!--       geom_point(shape=".") + -->
<!--         theme_bw() + -->
<!--         labs(title = .y, color = "Cell Type") + -->
<!--         ggrepel::geom_text_repel( -->
<!--           data= get_labels_clusters( -->
<!--             .x,  -->
<!--       			all_of(sample_column), -->
<!--       			umap_1,  -->
<!--       			umap_2 -->
<!--       			) , -->
<!-- 		aes(umap_1, umap_2, label = all_of(sample_column)), size = 2) + -->
<!--         guides(color = "none") +  -->
<!--         labs(title = .y) -->
<!--     })) |>  -->
<!--   mutate(overall_plot = map2(plot_by_doublet, plot_by_cell_type,  -->
<!--                              ~ .x + .y))  -->

<!-- plot_merged_combined_annotation_doublets <- merged_combined_annotation_doublets |>  -->
<!--                                             pull(overall_plot) |>  -->
<!--                                             wrap_plots(ncol = 1) + -->
<!--                                             plot_layout(guides = 'collect')   -->
<!--                                             # theme( -->
<!--                                             #   legend.position = "bottom",  -->
<!--                                             #   plot.margin = margin(10, 10, 10, 10, "cm") -->
<!--                                             # ) -->

<!-- # Print plot  -->
<!-- plot_merged_combined_annotation_doublets -->
<!--   # patchwork::wrap_elements() |>  -->
<!--   # map(~ .x |> -->
<!--   #        left_join( -->
<!--   #       params$x4 |> -->
<!--   #       purrr::reduce(bind_rows), by=".cell" -->
<!--   #     ) |> -->
<!--   #  -->
<!--   #     #join doublets identified -->
<!--   #     left_join( -->
<!--   #         params$x3 |> -->
<!--   #         purrr::reduce(bind_rows), by = c(".cell") -->
<!--   #       ) -->
<!--   # ) -->

<!-- ``` -->


```{r, echo = FALSE, warning=FALSE}
# Joining info and returning a list of tibbles
merged_combined_annotation_doublets <- list(
  print(length(calc_UMAP_dbl_report)),
  doublet_tbl,
  annotation_tbl
) |>
  pmap(
    ~ ..1 |>
        left_join(..2, by = ".cell") |>
        left_join(..3, by = ".cell")
  ) |> 
  enframe(name = "sample_id", value = "annotated_metadata") |> 
  mutate(
    sample_name = sample_names  # Using the sample_names list you already have
  ) |> 
  mutate(plot_by_doublet = map2(
    annotated_metadata,
    sample_name,
    ~ {
      # Sample to not overwhelm the plotting
      merged_combined_annotation_doublets = .x |> 
        nest(doublet_class = -scDblFinder.class) |>
        mutate(number_to_sample = if_else(scDblFinder.class == "singlet", 10000, Inf)) |>
        replace_na(list(number_to_sample = Inf)) |>
        unnest(doublet_class)

      # UMAP plot by doublet class
      merged_combined_annotation_doublets |> 
        ggplot(aes(umap_1, umap_2, color = scDblFinder.class)) +
        geom_point(shape = ".", size = 10) +
        theme_bw() +
        labs(title = .y, color = "Cell Type") + 
        ggrepel::geom_text_repel(
          data = get_labels_clusters(
            .x, 
            scDblFinder.class,
            umap_1, 
            umap_2
          ),
          aes(umap_1, umap_2, label = scDblFinder.class), size = 4
        ) +
        guides(color = "none") + 
        labs(title = .y)
    }
  )) |> 
  mutate(plot_by_cell_type = map2(
    annotated_metadata,
    sample_name,
    ~ {
      # Sample to not overwhelm the plotting
      merged_combined_annotation_doublets = .x |> 
        nest(doublet_class = -scDblFinder.class) |>
        mutate(number_to_sample = if_else(scDblFinder.class == "singlet", 10000, Inf)) |>
        replace_na(list(number_to_sample = Inf)) |>
        unnest(doublet_class)

      # UMAP plot by cell annotation
      merged_combined_annotation_doublets |> 
        ggplot(aes(umap_1, umap_2, color = !!sym(cell_ann_col))) +  # Using the cell annotation column dynamically
        geom_point(shape = ".") +
        theme_bw() +
        labs(title = .y, color = "Cell Type") +
        ggrepel::geom_text_repel(
          data = get_labels_clusters(
            .x, 
            !!sym(cell_ann_col),  # Assuming cell_ann_col contains your cell annotations
            umap_1, 
            umap_2
          ),
          aes(umap_1, umap_2, label = !!sym(cell_ann_col)), size = 2
        ) +
        guides(color = "none") + 
        labs(title = .y)
    }
  )) |> 
  mutate(overall_plot = map2(plot_by_doublet, plot_by_cell_type, 
                             ~ .x + .y)) 

# Combine the plots
plot_merged_combined_annotation_doublets <- merged_combined_annotation_doublets |> 
  pull(overall_plot) |> 
  wrap_plots(ncol = 1) +
  plot_layout(guides = 'collect')

# Print plot 
plot_merged_combined_annotation_doublets


```

## Singlet and Doublet Cell Distributions Across Tissues 

Each bar in the plot corresponds to a specific cell type within each tissue.

From this plot, we can infer: 

1. The overall quality of the cell separation process in the sequencing data, indicated by the proportion of singlets to doublets.
2. Potential differences in the rate of doublet formation between cell types, which might be related to cell size and tissue type

```{r, out.width='100%', fig.width=15, fig.height=10, warning=FALSE, message=FALSE, echo=FALSE}
# 2a) Create the composition of the doublets
doublet_composition<- merged_combined_annotation_doublets |>
  mutate(doublet_composition = map2(
    annotated_metadata,
    sample_name,
    ~ {
      #browser()
      .x|>
      dplyr::select(!!sym(sample_column), scDblFinder.class)}
      )) |>
      # table()|>
  dplyr::select(sample_name, doublet_composition) |>
  deframe()

  # merged_combined_annotation_doublets <-
  #   merged_combined_annotation_doublets |>
  # mutate(doublet_composition_plot = doublet_composition |>
  #          group_by(params$x5, all_of(params$x5)) |>
  #          mutate(proportion = count_class/sum(count_class)) |>
  #          ungroup()
  #        )

# calculate proportion and plot
#calculate proportion and plot
  merged_combined_annotation_doublets <-
    merged_combined_annotation_doublets |>
  mutate(doublet_composition_plot = map(
    annotated_metadata,
      ~ .x |>
      # browser() |>
  # create frequency column
  dplyr::count(.data[[sample_column]], .data[[cell_ann_col]], scDblFinder.class, name= "count_class") |>
    group_by(.data[[sample_column]], .data[[cell_ann_col]]) |>
    mutate(proportion = count_class/sum(count_class)) |>
    ungroup() |>

  # mutate(frequency = nCount_SCT/sum(nCount_SCT)*100) |>
  #
  # # create the proportion column
  # group_by(sample, scDblFinder.class) |>
  # mutate(tot_sample_proportion = sum(frequency)) |>
  # mutate(proportion = (frequency * 1)/tot_sample_proportion) |>

  #plot proportion
  ggplot(aes(x = .data[[cell_ann_col]] , y = proportion, fill = scDblFinder.class)) +
  geom_bar(stat = "identity") +
  theme_bw() +
  facet_wrap(~ orig.ident) +
  theme(axis.text.x=element_text(angle=70, hjust=1))
  ))

plot_merged_combined_annotation_doublets<- merged_combined_annotation_doublets|>
    pull(doublet_composition_plot)|>
    wrap_plots(ncol = 1) +
    plot_layout(guides = 'collect')

# Print plot
plot_merged_combined_annotation_doublets

```

<!-- ```{r, echo = FALSE} -->

<!-- # Doublet composition calculation using dynamic sample_column -->
<!-- doublet_composition <- merged_combined_annotation_doublets |> -->
<!--   mutate(doublet_composition = map2( -->
<!--     annotated_metadata, -->
<!--     sample_name, -->
<!--     ~ { -->
<!--       .x |>  -->
<!--       dplyr::select(!!sym(sample_column), scDblFinder.class)  # Using dynamic sample column -->
<!--     } -->
<!--   )) |> -->
<!--   dplyr::select(sample_name, doublet_composition) |> -->
<!--   deframe() -->

<!-- # Calculate proportions and plot using dynamic columns -->
<!-- merged_combined_annotation_doublets <-  -->
<!--   merged_combined_annotation_doublets |> -->
<!--   mutate(doublet_composition_plot = map( -->
<!--     annotated_metadata, -->
<!--     ~ { -->
<!--       .x |>  -->
<!--       # Create frequency column using dynamic sample_column and cell_ann_col -->
<!--       dplyr::count(!!sym(sample_column), !!sym(cell_ann_col), scDblFinder.class, name = "count_class") |>  -->
<!--       group_by(!!sym(sample_column), !!sym(cell_ann_col)) |>  -->
<!--       mutate(proportion = count_class / sum(count_class)) |>  -->
<!--       ungroup() |> -->

<!--       # Plot proportion using dynamic columns -->
<!--       ggplot(aes(x = !!sym(cell_ann_col), y = proportion, fill = scDblFinder.class)) + -->
<!--         geom_bar(stat = "identity") + -->
<!--         theme_bw() + -->
<!--         facet_wrap(as.formula(paste0("~ ", sample_column))) +  # Facet by dynamic sample column -->
<!--         theme(axis.text.x = element_text(angle = 70, hjust = 1)) -->
<!--     } -->
<!--   )) -->

<!-- # Combine plots -->
<!-- plot_merged_combined_annotation_doublets <- merged_combined_annotation_doublets |>  -->
<!--   pull(doublet_composition_plot) |>  -->
<!--   wrap_plots(ncol = 1) + -->
<!--   plot_layout(guides = 'collect') -->

<!-- # Print plot  -->
<!-- plot_merged_combined_annotation_doublets -->
<!-- ``` -->
