---
title: "Doublet identification report"
author: "SS"
date: "2023-12-05"
output: html_document
params: 
  x1: "NA"
  x2: "NA"
  x3: "NA"
  x4: "NA"
  x5: "NA"
  x6: "NA"
---

```{r setup, include=FALSE}
library(dplyr)
library(tidyr)
library(purrr)
# library(sccomp)
library(ggrepel)
library(Seurat)
library(glue)
library(scDblFinder)
library(Seurat)
library(tidyseurat)
library(tidySingleCellExperiment)
library(patchwork)
library(tibble)
library(scran)

get_labels_clusters = function(.data, label_column, dim1, dim2){

    tidy_dist = function(x1, x2, y1, y2){

        tibble(x1, x2, y1, y2) %>%
            rowwise() %>%
            mutate(dist = matrix(c(x1, x2, y1, y2), nrow = 2, byrow = T) %>% dist()) %>%
            pull(dist)

    }

    label_column = enquo(label_column)
    dim1 = enquo(dim1)
    dim2 = enquo(dim2)

    .data %>%
        nest(data = -!!label_column) %>%
        mutate(
            !!dim1 := map_dbl(data, ~ .x %>% pull(!!dim1) %>% median()),
            !!dim2 := map_dbl(data, ~ .x %>% pull(!!dim2) %>% median())
        ) %>%
        dplyr::select(-data)
}
```

## Comprehensive UMAP Visualization of Cell Typing and Doublet Detection Across Tissue Samples
- This visualization highlights the clustering of cell types and identifies singlets and doublets in the population. 
- This allows for an exploration of similarities and differences in gene expression profiles between cells from different tissues. 

```{r, out.width='100%', fig.width=15, fig.height=10, warning=FALSE, message=FALSE, echo=FALSE}
## Adjusting plot size according to the number of samples 
# num_samples <- length(params$x1$Tissue)
# 
# # Calculate the grid layout based on the number of samples
# num_columns <- ceiling(sqrt(num_samples))
# num_rows <- ceiling(num_samples / num_columns)


# Joining info and returning a list opf tibbles
merged_combined_annotation_doublets <- list(
  #params$x1,
  params$x2,
  params$x3,
  params$x4
) |>
    pmap(
        ~ ..1 |>
            left_join(..2, by = ".cell") |>
            left_join(..3, by = ".cell") 
          #left_join(..4, by = ".cell")
    ) |> 
  enframe(name = "sample_id", value = "annotated_metadata")|> 
  mutate(
    sample_name = map(annotated_metadata,  ~ .x |> pull(params$x5[[1]]))) |> 
  mutate(plot_by_doublet = map2(
    annotated_metadata,
    sample_name,
          ~ {
            #browser()
            merged_combined_annotation_doublets = .x |>
      
            #Sample to non overwhelmm the plotting
            nest(doublet_class = -scDblFinder.class) |>
            mutate(number_to_sample = if_else(scDblFinder.class=="singlet", 10000, Inf)) |>
            replace_na(list(number_to_sample = Inf)) |>
            # mutate(
            #   doublet_class = map2(doublet_class, number_to_sample, ~ .x |> sample_n(min(n(), .y)))
            #   )
            unnest(doublet_class) 
            # doublet_plots <- plot_by_doublet$doublet_class[[1]]|>
            
            merged_combined_annotation_doublets |> 
            ggplot(aes(umap_1, umap_2, color = scDblFinder.class)) +
              geom_point(shape=".", size = 10) +
              theme_bw() +
              labs(title = .y, color = "Cell Type") + 
              ggrepel::geom_text_repel(
                data= get_labels_clusters(
          			.x, 
          			scDblFinder.class,
          			umap_1, 
          			umap_2
          			) ,
      		aes(umap_1, umap_2, label = scDblFinder.class), size = 4) +
              guides(color = "none")+ 
        labs(title = .y)
          #print(plot_by_doublet)
           #return(plot_by_doublet)
          })) |> 
  
  mutate(plot_by_cell_type = map2(
    annotated_metadata,
    sample_name,
          ~ {
      #browser()
      merged_combined_annotation_doublets = .x |>
      # Sample to non overwhelmm the plotting
      nest(doublet_class = -scDblFinder.class) |>
      mutate(number_to_sample = if_else(scDblFinder.class=="singlet", 10000, Inf)) |>
        replace_na(list(number_to_sample = Inf)) |>
        #mutate(doublet_class = map2(doublet_class, number_to_sample, ~ .x |> sample_n(min(n(), .y))))|>
      unnest(doublet_class)
      
      merged_combined_annotation_doublets |> 
      ggplot(aes(umap_1, umap_2, color = all_of(params$x6))) +
      geom_point(shape=".") +
        theme_bw() +
        labs(title = .y, color = "Cell Type") +
        ggrepel::geom_text_repel(
          data= get_labels_clusters(
            .x, 
      			all_of(params$x5),
      			umap_1, 
      			umap_2
      			) ,
		aes(umap_1, umap_2, label = all_of(params$x5)), size = 2) +
        guides(color = "none") + 
        labs(title = .y)
    })) |> 
  mutate(overall_plot = map2(plot_by_doublet, plot_by_cell_type, 
                             ~ .x + .y)) 

plot_merged_combined_annotation_doublets <- merged_combined_annotation_doublets |> 
                                            pull(overall_plot) |> 
                                            wrap_plots(ncol = 1) +
                                            plot_layout(guides = 'collect')  
                                            # theme(
                                            #   legend.position = "bottom", 
                                            #   plot.margin = margin(10, 10, 10, 10, "cm")
                                            # )

# Print plot 
plot_merged_combined_annotation_doublets
  # patchwork::wrap_elements() |> 
  # map(~ .x |>
  #        left_join(
  #       params$x4 |>
  #       purrr::reduce(bind_rows), by=".cell"
  #     ) |>
  # 
  #     #join doublets identified
  #     left_join(
  #         params$x3 |>
  #         purrr::reduce(bind_rows), by = c(".cell")
  #       )
  # )

```

## Singlet and Doublet Cell Distributions Across Tissues 

Each bar in the plot corresponds to a specific cell type within each tissue.

From this plot, we can infer: 

1. The overall quality of the cell separation process in the sequencing data, indicated by the proportion of singlets to doublets.
2. Potential differences in the rate of doublet formation between cell types, which might be related to cell size and tissue type

```{r, out.width='100%', fig.width=15, fig.height=10, warning=FALSE, message=FALSE, echo=FALSE}
# 2a) Create the composition of the doublets

doublet_composition<- merged_combined_annotation_doublets |>
  mutate(doublet_composition = map2(
    annotated_metadata,
    sample_name,
    ~ {
      #browser() 
      .x|> 
      dplyr::select(all_of(params$x5), scDblFinder.class)}
      )) |>
      # table()|>
  dplyr::select(sample_name, doublet_composition) |>
  deframe()

  # merged_combined_annotation_doublets <-
  #   merged_combined_annotation_doublets |>
  # mutate(doublet_composition_plot = doublet_composition |> 
  #          group_by(params$x5, all_of(params$x5)) |>
  #          mutate(proportion = count_class/sum(count_class)) |>
  #          ungroup()
  #        )
  
# calculate proportion and plot
#calculate proportion and plot
  merged_combined_annotation_doublets <- 
    merged_combined_annotation_doublets |>
  mutate(doublet_composition_plot = map(
    annotated_metadata,
      ~ .x |> 
      # browser() |> 
  # create frequency column
  dplyr::count(.data[[params$x5]], .data[[params$x6]], scDblFinder.class, name= "count_class") |> 
    group_by(.data[[params$x5]], .data[[params$x6]]) |> 
    mutate(proportion = count_class/sum(count_class)) |>
    ungroup() |>
    
  # mutate(frequency = nCount_SCT/sum(nCount_SCT)*100) |>
  # 
  # # create the proportion column
  # group_by(sample, scDblFinder.class) |>
  # mutate(tot_sample_proportion = sum(frequency)) |>
  # mutate(proportion = (frequency * 1)/tot_sample_proportion) |>
  
  #plot proportion
  ggplot(aes(x = .data[[params$x6]] , y = proportion, fill = scDblFinder.class)) +
  geom_bar(stat = "identity") +
  theme_bw() +
  facet_wrap(~sampleName) +
  theme(axis.text.x=element_text(angle=70, hjust=1))
  )) 
  
  plot_merged_combined_annotation_doublets<- merged_combined_annotation_doublets|> 
    pull(doublet_composition_plot)|> 
    wrap_plots(ncol = 1) +
    plot_layout(guides = 'collect')  
 
# Print plot 
plot_merged_combined_annotation_doublets

```



