---
title: "pseudobulk analysis report"
author: "SS"
date: "2024-01-24"
output: html_document
params:
  x1: "NA"
  x2: "NA"
  x3: "NA"
---

```{r, echo=FALSE,results='hide', warning=FALSE, message=FALSE}
library(ggplot2)
library(stringr)
library(tidyverse)
library(tidybulk)
library(tidyseurat)
#library(tidysc)
library(tidyHeatmap)
library(purrr)
library(patchwork)
library(grid)
library(ComplexHeatmap)
library(ggrepel)
library(PCAtools)
library(tidySummarizedExperiment)
library(glue)
library(purrr)
library(plotly)
library(tidybulk)
#library(naniar) #NA 
library(magrittr)
library(here)
```

```{r, echo=FALSE,results='hide', warning=FALSE, message=FALSE}
# Load data
# pseudobulk <- do.call(cbind, params$x1)
pseudobulk <-  params$x1
#metadata_clinical_sample <- readRDS(params$metadata_path)
library(ggplot2)
# pseudobulk <-  params$x1
# Extract the proportion of variance explained by each principal component
# var_explained <- my_pca$sdev^2
# var_explained <- var_explained / sum(var_explained)
# cum_var_explained <- cumsum(var_explained)

# Find the number of components that explain at least 90% of the variance
# num_components <- which(cum_var_explained >= 0.9)[1]
```


```{r, echo=FALSE,results='hide', warning=FALSE, message=FALSE}
#pbmc_pseudobulk from sce:
pbmc_pseudobulk <- 
  pseudobulk %>% 
  # filter(data_source == assay) |> 
  #separate( .sample, c("single_cell_rna_id", "batch1"), "__" , remove=FALSE) |> 
  #left_join(metadata_clinical_sample |> tidybulk::pivot_sample(sample)) |>
  tidybulk::identify_abundant() %>%
  tidybulk::scale_abundance(method = "TMMwsp") 

data_for_pca = 
  pbmc_pseudobulk |> 
  keep_abundant() |> 
  keep_variable(.abundance = "count_scaled", top=500) |> 
  dplyr::select(-TMM, -multiplier, -count_scaled) |> 
  tidybulk::scale_abundance(method = "TMMwsp") 
```

## Checking that the input counts don't have global sequencing-depth effect

```{r, out.width='100%', fig.width=15, fig.height=10, warning=FALSE, message=FALSE, echo=FALSE}

data_for_pca |> 
  ggplot(aes(count_scaled + 1, color=.sample)) + geom_density(alpha=0.3) + scale_x_log10() + guides(color="none")
```

## Calculate PCA of pseudobulk

```{r, echo=FALSE,results='hide', warning=FALSE, message=FALSE}
metadata =
  data_for_pca |>
  pivot_sample() |>
  dplyr::select(any_of(params$x2), .sample, alive, any_of(params$x3), .aggregated_cells)

metadata = as.data.frame(metadata)
rownames(metadata) = metadata$`.sample`
# metadata = metadata[,-1]

my_pca = 
  data_for_pca@assays@data$count_scaled |> 
  log1p() |> 
  scale() |> 
  pca(metadata = metadata) 
#
# Extract the proportion of variance explained by each principal component
var_explained <- my_pca$sdev^2
var_explained <- var_explained / sum(var_explained)
cum_var_explained <- cumsum(var_explained)

# Find the number of components that explain at least 90% of the variance
num_components <- which(cum_var_explained >= 0.9)[1]
## Without metadata 
# my_pca =
#   data_for_pca@assays@data$count_scaled |>
#   log1p() |>
#   scale() |>
#   prcomp()

```

```{r, echo=FALSE,results='hide', warning=FALSE, message=FALSE}
# Find the number of components that explain at least 90% of the variance
num_components <- which(cum_var_explained >= 0.9)[1]
```

## Scree plot 
Graphical representation to show the proportion of variance explained by each principal component. 
This gives an idea of how many principal components we need to keep to represent the data faithfully. In this case we see a gradual decrease of variance explained, indicating that we might need up to principal component `num_components` for explaining 90% of the variance.

```{r, out.width='100%', fig.width=15, fig.height=10, warning=FALSE, message=FALSE, echo=FALSE}
library(ggplot2)

# # Extract the proportion of variance explained by each principal component
# var_explained <- my_pca$sdev^2
# var_explained <- var_explained / sum(var_explained)
# cum_var_explained <- cumsum(var_explained)

# Create a data frame for plotting
scree_data <- data.frame(PC = seq_along(var_explained), Variance = var_explained)

# Create the scree plot
ggplot(scree_data, aes(x = PC, y = Variance)) +
    geom_line() +
    geom_point() +
    theme_minimal() +
    labs(title = "Scree Plot", x = "Principal Component", y = "Proportion of Variance Explained")
```

## Principal Component Associations with Biological Variables
Here we see which variable is associated with which principal component. We hope the biological variable are associated with the top principal components.

In our sample data set we're clustering by Tissue type: Samples from the same tissue type cluster together in the PCA space, which indicates that the gene expression profiles are similar within a tissue type 

The distance of the points from the origin (where PC1 and PC2 both equal zero) indicates how much variance each sample has relative to the principal components. Samples that are further out along PC1 or PC2 axes have higher variance for those components.

```{r, out.width='100%', fig.width=15, fig.height=10, warning=FALSE, message=FALSE, echo=FALSE}
# x<- plot(my_pca$rotated[, "PC1"], my_pca$rotated[, "PC2"],
#      xlab = "PC1", ylab = "PC2", 
#      main = "PCA Plot",
#      asp = 1)
# x

x<- ggplot(my_pca$metadata, aes(x = my_pca$rotated[, "PC1"], y = my_pca$rotated[, "PC2"], color = my_pca$metadata[[params$x2]]))  +
  geom_point() +
  theme_minimal() +
  labs(title = "PCA Plot Colored by Tissue Type",
       x = "Principal Component 1",
       y = "Principal Component 2") +
  scale_color_discrete(name = "Tissue Type")
x
```

## Cell type clustering
- The separation or clustering of points with the same color might suggest that similar cell types have similar gene expression profiles, while different colors that group together could indicate distinct profiles between cell types. 
- The distance between the points on the plot reflects the similarity or dissimilarity in their gene expression data, as captured by the PCA.

```{r, out.width='100%',  warning=FALSE, message=FALSE, echo=FALSE}
data_for_pca |>
tidybulk::reduce_dimensions(method="PCA") |>
tidybulk::pivot_sample() |>
ggplot(aes(PC1, PC2, color=data_for_pca[[params$x3]])) +
geom_point() +
  theme_bw() +
  theme(
    legend.position = "right", # or choose "bottom" if you prefer
    legend.key.size = unit(0.2, "cm"), # Adjust the size of the legend keys
    legend.text = element_text(size = 3), # Adjust the text size in the legend
    legend.spacing.y = unit(0.1, "cm") # Adjust the spacing between legend entries
  )
```

<!-- Batch clustering -->

<!-- ```{r, out.width='100%',  warning=FALSE, message=FALSE, echo=FALSE} -->
<!-- data_for_pca |> -->
<!-- tidybulk::reduce_dimensions(method="PCA") |> -->
<!-- tidybulk::pivot_sample() |> -->
<!-- ggplot(aes(PC1, PC2, color=factor(Tissue))) + -->
<!-- geom_point() + -->
<!-- stat_ellipse() + -->
<!--   theme_bw()  -->
<!-- ``` -->