---
title: "Empty droplets"
author: "Elisa Roesti"
date: '2022-08-31'
params:
  reports_directory:
  file1: 
  file2:
output: md_document
---

```{r setup, include=FALSE}
library(DropletUtils)
library(tibble)
library(dplyr)
library(tidyr)
library(ggplot2)
library(tidyseurat)
library(glue)
library(purrr)
library(edgeR)
library(here)
```

```{r, warning=FALSE, message=FALSE, echo=FALSE, results='hide'}
# Create dir
params$reports_directory |> dir.create( showWarnings = FALSE, recursive = TRUE)

# Read files
demultiplexed = dir(params$file1, full.names = TRUE)
empty_droplet_identification_output = dir(params$file2, full.names = TRUE, pattern = "output.rds")

#read empty information
empty_df <- 
  empty_droplet_identification_output |>
  enframe(value = "file") |> 
  extract(file, "sample", "/([a-zA-Z0-9]+)__empty_droplet_identification_output.rds", remove = FALSE) |> 
  mutate(data = map(file, readRDS)) |> 
  unnest(data)



#define mito and ribo genes and add the plot:
demultiplexed_empty = demultiplexed[1] |> readRDS()

mito_genes <- grep("^MT-", rownames(demultiplexed_empty[["RNA"]]), value=T)
ribo_genes <-grep("^RP(S|L)", rownames(demultiplexed_empty[["RNA"]]), value=T)


# ma plot: mito is green, ribo is red
col <- rep('black',ncol(demultiplexed_empty))
col[rownames(demultiplexed_empty) %in% mito_genes] <-'green'
col[rownames(demultiplexed_empty) %in% ribo_genes] <-'red'
  
```


Visualize barcode rank plots for each sample

```{r, fig.height=10, fig.width=10,  warning=FALSE, message=FALSE, echo=FALSE}
#Code used to calculate the ranking (which is already present in our seurat), can be find at https://github.com/Melbourne-COVID-Predict/PBMC/blob/main/analysis/pseudobulk/pseudobulk_analysis_1.Rmd

#plot the graphs faceted:
  empty_df%>%
  sample_frac(0.1) %>% #between 0-1, 1=all cells
  ggplot2::ggplot(aes(rank, total)) + 
  geom_point() + 
  facet_wrap(~sample, nrow=4) + 
  geom_line(aes(rank, fitted), color="red") + 
  geom_hline(aes(yintercept = knee), color="dodgerblue") +
  geom_hline(aes(yintercept = inflection), color="forestgreen") +
  scale_x_log10() + 
  scale_y_log10() + 
  theme_bw() +
  theme(aspect.ratio=1) +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1))

```


```{r}
empty_df  |> select(empty_droplet) |> table()

```



Number of non-empty droplets, everything above knee is retained.

```{r,  warning=FALSE, message=FALSE, echo=FALSE}
# Number of non-empty droplets -------------------------------------------------
summary(empty_df$FDR < 0.001)
```

```{r,  warning=FALSE, message=FALSE, echo=FALSE}
is.cell <- empty_df$FDR <= 0.001
table(is.cell) #true 81090 cells, 8429 false
```

Check if p-values are lower-bounded by 'niters' (increase 'niters' if any Limited==TRUE and Sig==FALSE), with niters being: An integer scalar specifying the number of iterations to use for the Monte Carlo p-value calculations.

```{r,warning=FALSE, message=FALSE, echo=FALSE}
table(Sig=is.cell, Limited=empty_df$Limited)
```


MA plot (only if empty droplets have been identified)

```{r, fig.height=10, fig.width=10, warning=FALSE, message=FALSE, echo=FALSE}
if(empty_df |> filter(empty_droplet) |> nrow() != 0)
  tibble(
  input_files = demultiplexed,
  empty_droplets = empty_droplet_identification_output
) |> 
  mutate(MA_plot = map2(
    input_files, empty_droplets,
    ~ {
      
      demultiplexed_empty = readRDS(here(.x))
      empty_df = readRDS(here(.y))
      
      extra <- empty_df |> filter(FDR < 0.001) |> pull(.cell)
      amb  <- empty_df |> filter(is.na(FDR)) |> pull(.cell)
      extra.mu <- demultiplexed_empty |> filter(.cell %in% extra) |> GetAssayData("RNA",slot = "counts") |> rowSums()
      amb.mu  <-demultiplexed_empty |> filter(.cell %in% amb) |> GetAssayData("RNA",slot = "counts") |> rowSums()
      

      edgeR::maPlot(extra.mu, amb.mu, normalize = TRUE, col = col, plot.it = FALSE) |> 
        
      # ggplot
      enframe() |> 
        pivot_wider() |> 
        unnest(A, M, w, v) |> 
        ggplot(aes(A, M, color = col, size = col)) + 
        geom_point() + 
        scale_color_manual(values = c("black", "red", "blue")) + 
        scale_size_manual(values = c(0.2, 2, 2)) +
        theme_bw()


    }
  )) |> 
  pull(MA_plot)

```

(mitochondria genes in green, ribosomal genes in red)



Histogram of p-values: (only if empty droplets have been identified)

```{r,  warning=FALSE, message=FALSE, echo=FALSE}
if(empty_df |> filter(empty_droplet) |> nrow() != 0){
  q = quantile(empty_df$Total, 0.1)

hist(empty_df$PValue[empty_df$Total <= q & empty_df$Total > 0],
     main = glue("Droplets with 0 < libsize <= {q}"),
     xlab = "P-value")
}

```

